\chapter{Introduction}

Crisis Text Line (CTL) \cite{ctl} is an organization that provides counseling services to
young people in crisis through texting. The goal of this thesis is to supply CTL
counselors with assistive tools in order to offer their clients the best possible service.
This section explains the motivation behind this research project, the problems we
want to solve, the approaches to implement and evaluate, and the contributions made.

\section{Motivations for Mental Health Visualizations}

The main motivating factor for this thesis is to help people with mental health crises.
Many people suffer from depression, suicidal thoughts, and emotional stress every
day. Crisis Text Line provides an outlet for clients to discuss their issues and ask for
support. However, there is a shortage of counselors compared to the number of people
seeking aid. Each counselor may have to manage various conversations continuously
for several hours. Counselors also have to spend time on extraneous things such as
reports. In addition to maximizing counseling time for clients, we want to assist
counselors because crisis counselors tend to have a high attrition rate due to burnout
and low morale. Many of these counselors are simply volunteers who undergo a short
period of training. These constraints motivate us to maximize the amount of time
and brainpower counselors spend on client support. Time is a critical factor in mental
health situations because clients may be at risk of suicide or physical harm.

Fortunately, the unique thing about a texting hotline is that the use of written
communication allows computer programs to analyze and extract meaningful data
from the text. Topic modeling is a machine learning technique that discovers abstract
topics occurring in a set of documents. It can be useful for summarizing large amounts
of text. Counselors may benefit from summaries of their various conversations with
clients, but topic modeling is a complex and advanced artificial intelligence concept.
Therefore, we would like to provide counselors with an easy method of understanding
the data through visualizations. Visualization is a powerful approach for presenting
data that most people are familiar with.

\section{Problem Definition}

As mentioned in the motivation section, some main difficulties with Crisis Text Line
are that there are not enough counselors to talk with all of the clients in crisis, the
counselors are usually context switching between multiple different client texts at a
time, they spend a nontrivial amount of time on other necessary tasks, and counselors
may feel burned out. Since it is more difficult to control external factors such as the
number of available counselors, we focus on tackling two specific problems:
\begin{enumerate}
  \item Reduce the amount of time counselors spend not talking to clients.
  \item Reduce the cognitive load of counselors so they feel less burned out.
\end{enumerate}

\section{Hypotheses}

We believe a variety of topic model visualizations will offer assistance in solving the
proposed problems. Using topic models, mental health conversations may be
summarized by a combination of topics, such as job-related issues, family troubles,
relationship difficulties, or self-injurious behavior, to name a few. Topic models also
provide indexing information, which tells us where each specific topic can be found
in the conversation text.

\subsection{Context Switching}

As the medium of texting usually involves gaps in response time during a conversation,
counselors often switch context between talking to different clients. When a counselor
returns to a previous conversation after a client response, he or she may have to spend
time recalling what that particular conversation was about. However, if the counselor
was given a visual summary of the conversation, with the option of quickly reading
through chat details, less time may be spent recognizing the conversation topics. This
approach can minimize both the time a counselor spends not talking to a client and
the cognitive load that context switching has on the counselor.

\subsection{Conversation Trends}

Another tactic to aid the counselor's memory management of different conversations
may be to display topic trends over time for each conversation. This level of detail
could remind counselors of the topics most recently discussed as well as the previous
talking points. Showing trends, including where topics appear in the course of a
conversation and how they accumulate, may also improve the quality of client service.
A chart of topic trends could alert the counselor to important focus points. For
example, if the topic of self-injurious behavior is on the rise, the counselor might
want to react in a certain way to prevent escalation of injury. Conversation trends
can also be useful for organization leaders to detect patterns that might be of use in
supporting clients.

\subsection{Shift Changes and Repeat Clients}

Another situation that could potentially benefit from assistance is shift changes.
Counselors usually handle incoming client texts in shifts. A shift change may
occur in the middle of a conversation, in which case the leaving counselor gives the
incoming counselor a brief summary of the talking points so he or she can take over.
However, this summary is general and transient, and the incoming counselor would
have to take time scanning through the existing conversation text for details.
Similarly, counselors have the option of reading past conversations for a client that has
texted the hotline previously to gain context. This is a useful but time-consuming
process when taking on a repeat client. We suspect that a permanent visual summary
computed using topic models would be more helpful for the counselor in both
situations. The visualizations can provide different levels of detail depending on what
the counselor needs to know about the conversation history. Visual indexing can
quickly point him or her to the parts of the conversation related to a certain topic.
This technique minimizes the amount of time necessary to search through the text
for details and potentially improves the quality of client service by better preparing
the new counselor for the interaction.

\subsection{Automated Reports}

Finally, certain crisis organizations require their counselors to complete reports on
their conversations with clients. Although these reports may be helpful for some,
they can be time-consuming to fill out manually. That time could be better spent
interacting with clients. Given that we can use algorithms to analyze conversation text
and extract information, we believe that this information can be used to automatically
pre-populate reports. A visual summary can even act as a more effective report on
its own, saving the counselors time by not having to fill out a report themselves.

\section{Contributions}

Based on a topic model developed from a collection of real mental health
conversations, I designed and implemented a website prototype for Crisis Text Line with four visualizations. These visualizations were designed based on four different levels of
granularity, so the counselor can choose the amount of detail he or she wants.

The \textbf{Topic List} visualization lists the topics discovered in a conversation that
are above a certain threshold. Topics are ordered from highest to lowest percentage
detected in the conversation. This visualization is a quick, glance-able summary of
the conversation topics.

The \textbf{Donut Chart} visualization adds a small level of detail by displaying the topic
proportions in a pie chart variation to show the parts of the whole relationship. User
interaction by hovering over the chart or the legend provides the topic percentages 
or quantitative information.

The \textbf{Line Chart} visualization reaches finer granularity by revealing topic
proportions at the message level, where each client message in the conversation is analyzed
for topics. A line exists for each topic above a certain threshold that shows the trend
of that topic throughout the conversation timeline. There is also the option of viewing
the accumulation of topic proportions across the conversation. When the user clicks
on a topic, points are displayed to reveal the client messages in the conversation that
contain the topic.

The \textbf{Scatter Plot} visualization is the deepest detail level, allowing the user to
click on the topic instances that occur throughout the conversation. The conversation
text then automatically scrolls to the appropriate message. The size of the scatter
plot points represent the proportion of that topic in the corresponding message.

\section{Thesis Outline}

Chapter two presents related work, consisting of topic models, topic model
visualizations for other fields, and mental health topic modeling, as well as the scope and
limitations of the thesis.

Chapter three describes the contextual inquiry done with crisis counselors to
analyze the specific needs of our users for a better design.

Chapter four discusses the design of the four visualizations contributed in this thesis:
a topic list, a donut chart, a multi-series line chart, and a scatter plot.

Chapter five explains how the system was implemented and lists the existing
technologies that were used.

Chapter six evaluates the visualizations based on user test results.

Chapter seven explores ideas for future work.

Finally, chapter eight discusses the main contributions presented in this thesis.
